### 2021-04-01 referee_response2.md

> 5. In the conclusion section, the formulation of the conclusions on section 3.2.2. are much stronger than justified: you say that you cannot completely rule out the manifold theory, even though you admit that the proxy you use is not convincing. Since hurried readers could read this conclusion section without reading the bulk of the paper, you leave them with the impression that you almost rule out the Manifold theory, while this is not the case, since your proxy for bar strength is very poor, and your sample is small. Please harmonize your conclusions with the reality you describe in the main section.

We have further toned down the conclusion on Manifold theory and believe it now includes most of the caveats given in the main section.

> 6. In the conclusions, page 11 left side lines 48-49, you suggest that the absence of galaxies with low pitch angles is not a selection effect. How does this square with the data of Kennicutt (1981), where nearly all Sa galaxies, as classified in the RSA catalogue, have pitch angles less than 10 degrees? Are you sure that you are not missing the tightly wound early type spiral galaxies?

Our sample selection is quite different to Kennicutt (1981) which may be one factor. It seems possibly that very tightly wound spirals may be more likely to be misidentified as smooth galaxies, and therefore not make it into the spiral sample in Galaxy Zoo at all. In addition a colleague compared Kennicutt (1981) pitch angles to those measured using the Spacfire software, and found that Sparcfire measured more open arms in general for the early type spirals. We have not run the Kennicutt sample through Galaxy Builder, so we cannot reliably comment if that would be an effect here, but will make sure to do that for the next phase of the project.

### 2021-04-05 referee_response2.md

> Please refrain from reformulating the Hubble classification to a scheme based on your own criteria: e.g. the prototype flocculent spiral galaxy NGC 2841 has condensations in its disc, but no clear spiral arms.

We have made edits to use the recommended language.

> 2. In this context, it is striking that the authors now choose NGC 4689 as a example of a flocculent galaxy, even though spiral structure is visible, in particular the eastern arm. True, Elmegreen & Elmegreen (1987) list it as a class 3 spiral, i.e. in the same category as NGC 2841, but the spiral arms are clearly there in the 3.6 micron Spitzer image. I suggest using NGC 2841 as a good example for a flocculent galaxy, and note that for NGC 5055 there is a paper by Michele Thornley -  1996ApJ...469L..45T - pointing out that in the K'-band there is spiral structure seen, even though it is also listed as a prototype flocculent spiral in Elmegreen & Elmegreen (1987).

We have edited to use NGC 2841 as the example of a flocculent galaxy.

@@

> 5. In the conclusion section, the formulation of the conclusions on section 3.2.2. are much stronger than justified: you say that you cannot completely rule out the manifold theory, even though you admit that the proxy you use is not convincing. Since hurried readers could read this conclusion section without reading the bulk of the paper, you leave them with the impression that you almost rule out the Manifold theory, while this is not the case, since your proxy for bar strength is very poor, and your sample is small. Please harmonize your conclusions with the reality you describe in the main section.

We have further toned down the conclusion on Manifold theory and believe it now includes most of the caveats given in the main section.

> 6. In the conclusions, page 11 left side lines 48-49, you suggest that the absence of galaxies with low pitch angles is not a selection effect. How does this square with the data of Kennicutt (1981), where nearly all Sa galaxies, as classified in the RSA catalogue, have pitch angles less than 10 degrees? Are you sure that you are not missing the tightly wound early type spiral galaxies?

Our sample selection is quite different to Kennicutt (1981) which we suspect is the primary factor. For example the most distant of the Kennicutt galaxies is at a similar redshift the lower limit of our sample. This means the mass range covered will be very different too, with the Kennicutt sample likely weighted to lower mass galaxies. Another factor, is that we have found that very tightly wound spirals may be misidentified as smooth galaxies at larger distances, and therefore not make it into the spiral sample in Galaxy Zoo at all. In addition a co-author (BK) compared Kennicutt (1981) pitch angles to those measured using the Spacfire software, and found that Sparcfire measured more open arms in general for the early type spirals. We have not run the Kennicutt sample through Galaxy Builder, so we cannot reliably comment if that would be an effect here, but will make sure to do that for the next phase of the project.

 We have added some additional lines about sample selection.

@@

> I did a cursory inspection of the early type spirals in the Yu & Ho (2018) sample, and find that most of the early type spirals (S0/a, Sa and Sab) are actually barred spirals when checking the classification in Hyperleda and displaying them in the legacysurvey viewer. Perhaps the authors can check the distribution of the results of Kennicutt (1981) in a diagram such as Figure 9, to see whether his sample is in agreement with the ideas of Pringle & Dobbs (2019)?

 We did this, and the Kennicutt sample is peaked at lower pitch angle, so the agreement is not as good. We have added a note about this to the paper, including comments on sample selection, and we include the plot for the referee.

### 2021-04-05 sections/summary.tex
Final edits from KLM (maybe)

We do not find any link between {\bf our measure of} bar strength and pitch angle in our sample. However, rather that a direct measure of bar strength, we make use of an available parameter which correlates with bar strength, so at best this observation {\bf is suggestive} that the primary mechanism driving the evolution of the spirals in our sample {\bf may not be} Manifold theory (see Section \ref{section:morphology-comparison-bar}), \textbf{however, since this is not the measure of bar strength predicted to correlate with pitch angle by \citet{2009MNRAS.400.1706A}, and those authors caution that the details of the bar strength measure can wash out the predicted correlation, this is not strong evidence against Manifold theory models}.

@@

In this work, we assume that spiral arms are equally likely to be identified and recovered at all pitch angles, which suggests the absence of galaxies at low pitch angles is not a selection effect. This is not an unfair assumption given the amount of human effort that went into obtaining spiral arm measurements (more so than any other pitch angle measurement method, with each galaxy receiving at least 30 human classifications). The galaxy sample used is not selected to be representative of the general spiral population, but is comparable in size to those used in other similar studies (\citealt{2013MNRAS.436.1074S}, \citealt{2019ApJ...871..194Y}, \citealt{2019arXiv190910291P}) and covers a range of masses ($9.45 < \log(M_* / M_\odot) < 11.05$) and spiral types.

>>>>

We do not find any link between {\bf our measure of} bar strength and pitch angle in our sample. However, rather that a direct measure of bar strength, we make use of an available parameter which correlates with bar strength, so at best this observation {\bf is suggestive} that the primary mechanism driving the evolution of the spirals in our sample {\bf may not be} Manifold theory (see Section \ref{section:morphology-comparison-bar}). \textbf{Since this is not the measure of bar strength predicted to correlate with pitch angle by \citet{2009MNRAS.400.1706A}, and those authors caution that the details of the bar strength measure can wash out the predicted correlation, this is not strong evidence against Manifold theory models}.

@@

In this work, we assume that spiral arms are equally likely to be identified and recovered at all pitch angles, which suggests the absence of galaxies at low pitch angles is not {\bf due to an inability of us to measure such arms}. This is not an unfair assumption given the amount of human effort that went into obtaining spiral arm measurements (more so than any other pitch angle measurement method, with each galaxy receiving at least 30 human classifications). The galaxy sample used is {\bf a random subset of a volume limited sample (see Figure \ref{fig:stellarmass}), and} is comparable in size to those used in other similar studies (\citealt{2013MNRAS.436.1074S}, \citealt{2019ApJ...871..194Y}, \citealt{2019arXiv190910291P}). The sample covers a range of masses ($9.45 < \log(M_* / M_\odot) < 11.05$) and spiral types, {\bf however it is possible that tightly wound spirals are preferentially missed by the pre-selection from {\it Galaxy Zoo}, if they are less obviously identified as having spirals at these distances ($0.02<z<0.055$). The \citet{1981AJ.....86.1847K} sample of 113 much more nearby spirals (all at $z<0.019$ and most at $z<0.009$), includes several with arms at much lower pitch angles, but very few arms which are loosely wound (none with $\cot \phi > 31^\circ$), meaning it does not match the same $\cot \phi$ constant model well, although as at the time the sample was set by data availability, it is unclear how conclusive this is. In a future version of {\it Galaxy Builder} we intend to include this sample as a comparison set.}
